A hidden Active Galactic Nuclei population: the first radio luminosity functions by process

### 2024-08-23 src/tex/ms.tex
[showyourwork] overleaf sync

Thanks to the added collecting area of the international stations, which increases the sensitivity of the array, the 0.\sarc\ 3 resolution images are \textit{deeper} than the 6\sarc\ \ resolution images made with similar or larger amounts of data. The derived galaxy properties from \citetalias{best_lofar_2023} (see \S~\ref{subsec:dr1}) were for LoTSS Deep Fields DR1, and thus we have to limit the source catalogues we wish to use. For both Lockman Hole and ELAIS-N1 catalogues derived from the 0.\sarc\ 3 resolution images, we first find the local rms in the companion 6\sarc\ \ resolution noise maps\footnote{\label{dr1}Available at \href{https://lofar-surveys.org/deepfields.html}{https://lofar-surveys.org/deepfields.html}} at the position of each source. Using the peak brightness of the source in the catalogue from the 0.\sarc\ 3 resolution image, we calculate the peak-to-noise ratio and only select sources which would be detectable at the 5$\sigma$ limit. This ensures we do not have sources in the 0.\sarc\ 3 resolution catalogue which do not have a match in DR1. We then cross-match based on radio position, to provide the radio information from the catalogues derived from both the 0.\sarc\ 3 resolution image and the 6\sarc\ \ resolution image. We trim the area down for each field using their respective optical starmasks\textsuperscript{\ref{dr1}}, by selecting sources which are not contained in the null areas of these masks. This removes sources where the multi-wavelength information is impacted by bright stars. Finally, we trim the area down to the 2.5 deg$\times$2.5 deg area which is covered by the 0.\sarc\ 3 resolution image. We are left with \variable{output/lockman_detectable.txt} sources in Lockman Hole, and \variable{output/en1_detectable.txt} sources in ELAIS-N1. We use the source name from the catalogue derived from the 6\sarc\ \ resolution to cross-match to the catalogue from \citetalias{best_lofar_2023}.

@@

\subsection{Separation of AGN activity and SF}
Once the AGN activity is identified from the 0.\sarc\ 3 resolution images, it is possible to separate it from star formation using the total flux density from the 6\sarc\ \ resolution images. We refer the reader to \S~5 in \citetalias{morabito_identifying_2022} for more details but outline the process here, which is depicted in Fig.~\ref{fig:flowchart}. Sources which are identified as radio excess according to the definition in \citetalias{best_lofar_2023} automatically have their total flux density from the 6\sarc\ resolution images added to the AGN category. For sources that are not radio excess, we attribute the total flux density from the 6\sarc\ resolution images to SF if in the 0.\sarc\ 3 resolution image they are either resolved or have no detection. Sources which do not meet these criteria are checked and if their $T_b$ is not above the limit for SF their 6\sarc\ resolution total flux density is also added to the SF category. The total flux density of any source in the 0.\sarc\ 3 resolution image which is $T_b$-identified as an AGN is assumed to be AGN flux density; this is subtracted from the total flux density measured from the 6\sarc\ \ resolution image, and what remains is assumed to be flux density from SF. \hl{how many sources per category}

@@

We follow the same method of constructing RLFs as in \cite{kondapally_cosmic_2022} and \cite{cochrane_lofar_2023}. As the 0.\sarc\ \ 3 resolution image is smaller than the 6\sarc\ resolution image, we first reproduce the published RLFs (Fig.~\ref{fig:rlfs}, left panel) using the galaxy classifications from \citetalias{best_lofar_2023} to check that we have implemented the method appropriately, and that the smaller area yields the same results. We use the completeness corrections from the respective papers. The only correction we do not implement is the photo-$z$ correction used in \cite{cochrane_lofar_2023}. The authors note that this correction is negligible, and as it is not implemented in \cite{kondapally_cosmic_2022} we chose to keep the method for RLF construction in this paper consistent across both samples. We use the same redshift range, $0.003 < z < 0.3$ and see that we indeed have good agreement with the published RLFs.

>>>>

Thanks to the added collecting area of the international stations, which increases the sensitivity of the array, the 0.\sarc\ 3 resolution images are \textit{deeper} than the 6\sarc\ \ resolution images made with similar or larger amounts of data. The derived galaxy properties from \citetalias{best_lofar_2023} (see \S~\ref{subsec:dr1}) were for LoTSS Deep Fields DR1, and thus we have to limit the source catalogues we wish to use. For both Lockman Hole and ELAIS-N1 catalogues derived from the 0.\sarc\ 3 resolution images, we first find the local rms in the companion 6\sarc\ \ resolution noise maps\footnote{\label{dr1}Available at \href{https://lofar-surveys.org/deepfields.html}{https://lofar-surveys.org/deepfields.html}} at the position of each source. Using the peak brightness of the source in the catalogue from the 0.\sarc\ 3 resolution image, we calculate the peak-to-noise ratio and only select sources which would be detectable at the 5$\sigma$ limit. This ensures we do not have sources in the 0.\sarc\ 3 resolution catalogue which do not have a match in DR1. We then cross-match based on radio position, to provide the radio information from the catalogues derived from both the 0.\sarc\ 3 resolution image and the 6\sarc\ \ resolution image. We trim the area down for each field using their respective optical starmasks\textsuperscript{\ref{dr1}}, by selecting sources which are not contained in the null areas of these masks. This removes sources where the multi-wavelength information is impacted by bright stars. Finally, we trim the area down to the 2.5 deg$\times$2.5 deg area which is covered by the 0.\sarc\ 3 resolution image. We are left with \variable{output/lockman_detectable.txt}sources in Lockman Hole, and \variable{output/en1_detectable.txt}sources in ELAIS-N1. We use the source name from the catalogue derived from the 6\sarc\ \ resolution to cross-match to the catalogue from \citetalias{best_lofar_2023}.

@@

\subsection{Separation of AGN activity and SF}
Once the AGN activity is identified from the 0.\sarc\ 3 resolution images, it is possible to separate it from star formation using the total flux density from the 6\sarc\ \ resolution images. We refer the reader to \S~5 in \citetalias{morabito_identifying_2022} for more details but outline the process here, which is depicted in Fig.~\ref{fig:flowchart}. Sources which are identified as radio excess according to the definition in \citetalias{best_lofar_2023} automatically have their total flux density from the 6\sarc\ resolution images added to the AGN category. For sources that are not radio excess, we attribute the total flux density from the 6\sarc\ resolution images to SF if in the 0.\sarc\ 3 resolution image they are either resolved or have no detection. Sources which do not meet these criteria are checked and if their $T_b$ is not above the limit for SF their 6\sarc\ resolution total flux density is also added to the SF category. The total flux density of any source in the 0.\sarc\ 3 resolution image which is $T_b$-identified as an AGN is assumed to be AGN flux density; this is subtracted from the total flux density measured from the 6\sarc\ \ resolution image, and what remains is assumed to be flux density from SF. \variable{output/flowchart_numbers.txt}

@@

We follow the same method of constructing RLFs as in \cite{kondapally_cosmic_2022} and \cite{cochrane_lofar_2023}. The 0.\sarc\ 3 resolution image is smaller than the 6\sarc\ resolution image, so we first reproduce the published RLFs (Fig.~\ref{fig:rlfs}, left panel) using the galaxy classifications from \citetalias{best_lofar_2023} to check that we have implemented the method appropriately, and that the smaller area yields the same results. We use the completeness corrections from the respective papers. The only correction we do not implement is the photo-$z$ correction used in \cite{cochrane_lofar_2023}. The authors note that this correction is negligible, and as it is not implemented in \cite{kondapally_cosmic_2022} we chose to keep the method for RLF construction in this paper consistent across both samples. We use the same redshift range, $0.003 < z < 0.3$ and see that we indeed have good agreement with the published RLFs.
